---
title: "Bagging"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bagging}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(SimpleEnsembleGroup17)
```

## Introduction

The `bagging` function from the `SimpleEnsembleGroup17` package provides a convenient way to train various types of models, including lasso regression, elastic net regression, and ridge models. In this vignette, we'll demonstrate how to use `bagging` with the Leukemia dataset from the [Source Url](https://www.ams.sunysb.edu/~pfkuan/Teaching/AMS597/Data/leukemiaDataSet.txt) and explain the different model types and their applications.

## The bagging Function

```{r bagging, eval=FALSE}
bagging <- function(X, y, model_type = "lasso", R=42) {
...
}
```

The `bagging` function has the following signature:

- `X`: This is the matrix or data frame containing the predictor variables (features)
- `y`: This is the target column in the dataset
- `model_type`: This parameter specifies the type of model to be trained. The default value is `"lasso"`, but it can be set to other values like  `"ridge"`, `"elastic_net"`. The code does not show outputs for `"logistic"`, `"random_forest"` or `"linear"`
- `R` : The R states the solution to sample with R times


## Load the Leukemia Dataset

```{r load-data}
url <- "https://www.ams.sunysb.edu/~pfkuan/Teaching/AMS597/Data/leukemiaDataSet.txt"
data <- read.table(url, header = TRUE)
data$Group <- ifelse(data$Group=="AML",1,0)
```

The Leukemia dataset contains information about "Gene expression", data provide information about which genes are active or inactive in a cell, and studying these patterns can help identify genes that are involved in leukemia development and progression. Our target variable consist of ALL/AML.


## Bagging
The Bagging technique is also a bootstrapping technique is like an ensembling technique which boosts accuracy and improves stability of a model. Lets see an example of Bagging on the Leukemia dataset.

```{r Bagging}
X <- data[, -which(names(data) == "Group")]
y <- data$Group
result<-bagging(X,y,model_type="lasso",R=42)
print(result$naive_importance)
```

In the above example, we train a bagging model to predict the Group of the Leukemia dataset. The parameters used are:

- `y = data$Group`: This is the target variable of the dataset
- `X =  data[, -which(names(data) == "Group")]` : This is the matrix containing the predictor variables, excluding the target variable.
- `model_type = "lasso"`: This parameter specifies that we want to train a lasso for bagging model.

## Conclusion

In this vignette, we demonstrated how to use the `bagging` function to train linear regression,logistic regression, ridge regression, elastic_net, lasso regression, and random forest models using the Leukemia dataset. The `bagging` function provides a convenient way for tasks such as handling categorical variables, missing data, sampling R times and model-specific parameter tuning.

By leveraging the `bagging` function, users can easily experiment with different model types and select the most appropriate one with r times for their specific problem and dataset. 
