---
title: "TopK"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TopK}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SimpleEnsembleGroup17)
```
## Introduction
The `topk_selector` function from the `SimpleEnsembleGroup17` package provides a convenient way for selecting the top K predictors from a dataset using various techniques such as correlation, t-test, and lasso regression. In this vignette, we'll demonstrate how to use `topk_selector` with the Leukemia dataset from the [Source Url](https://www.ams.sunysb.edu/~pfkuan/Teaching/AMS597/Data/leukemiaDataSet.txt) and explain the different model types and their applications.
```{r}
topk_selector <- function(X, y, technique = "correlation", method = "pearson", threshold = 0.05, top_k = 2) {
  # Function body...
}
```
The function has the following parameters:

- X: Predictor variables (features) in the dataset.
- y: Target variable (response) in the dataset.
- technique: The technique used for predictor selection. Options include "correlation", "t-test", and "lasso".
- method: The method used for correlation calculation (only applicable if technique is "correlation"). Default is "pearson".
- threshold: The threshold value for selecting predictors based on the specified technique. Default is 0.05.
- top_k: The number of top predictors to select. Default is 2.

## Load the Leukemia Dataset
```{r load-data}
url <- "https://www.ams.sunysb.edu/~pfkuan/Teaching/AMS597/Data/leukemiaDataSet.txt"
data <- read.table(url, header = TRUE)
data$Group <- ifelse(data$Group=="AML",1,0)
```
The Leukemia dataset contains information about "Gene expression", data provide information about which genes are active or inactive in a cell, and studying these patterns can help identify genes that are involved in leukemia development and progression. Our target variable consist of ALL/AML.


```{r}
X <- data[, -which(names(data) == "Group")]
y <- data$Group
result <- SimpleEnsembleGroup17::topk_selector(X, y, technique = "correlation", method = "pearson", threshold = 0.05, top_k = 5)
```

## Conclusion
The topk_selector function provides a flexible way to select the top predictors from a dataset using different techniques. By adjusting the parameters such as technique, method, threshold, and top_k, users can tailor the predictor selection process according to their specific needs and preferences.
